# Results

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r}

library(cpsR)
library(tidyverse)
library(ggplot2)
library(DataExplorer)
library(data.table)

df <- get_asec(year = 2023, vars = c("PERIDNUM", "A_FNLWGT", "GESTFIPS", "GTCO", "HEA", "A_AGE", "A_SEX", "PRCITSHP","PRDTRACE", "PEMLR","A_HGA","PEARNVAL","DIS_HP","NOW_COV", "MOOP", "PTOTVAL"), key="033d6fdb469d970444d91cda781be61fa0d46099", tibble = FALSE)


# We rename each variable as follows:
df <- df %>% 
  rename(
    ID = "peridnum",
    Wt = "a_fnlwgt",
    State = "gestfips",
    Health = "hea",
    Age = "a_age",
    Sex = "a_sex",
    Citz = "prcitshp",
    Race = "prdtrace",
    Empt = "pemlr",
    Edu = "a_hga",
    Earn = "pearnval",
    Pwd = "dis_hp",
    Ins = "now_cov",
    MExp = "moop",
    Inc = "ptotval"
    )

# Since there are certain features whose categories are too detailed for our purposes, we make the following preprocessing:
  
### Sex ###
df$Sex <- factor(df$Sex, levels = c(1, 2), labels = c("Male", "Female"))

### Citizenship ###
df$Citz <- factor(df$Citz, levels = c(1, 2, 3, 4, 5), labels = c("Native", "Native", "Native", "Naturalized", "Not US Citizen"))
# We group:
#     categories 1 (Native, born in the US), 2 (Native, born in PR or US outlying are) and 3 (Native, born abroad of US parent(s)) into "Native". 
#table(df$Citz)

### Employment ###
df$Empt <- factor(df$Empt, levels = c(0, 1, 2, 3, 4, 5, 6, 7), labels = c("NIU","Employed", "Employed", "Unemployed", "Unemployed", "Retired", "Disabled", "Other"))
# We group: 
#     categories 1 (Employed - at work) and 2 (Employed - absent) into "Employed"
#     categories 3 (Unemployed - on layoff) and 4 (Unemployed - looking) into "Unemployed". 
#table(df$Empt)

### Education ###
#table(df$Edu)
df$Edu <- ifelse(df$Edu >= 31 & df$Edu <= 38, 30, df$Edu)
# We group categories 31 to 38 into 30 (than we'll label it as "High school Incomplete")
df$Edu <- factor(df$Edu, levels = c(0, 30, 39, 40, 41, 42, 43, 44, 45, 46), labels = c("Children","H.S. Inc.", "H.S. Compl.", "College Inc.", "Associate", "Associate", "Bachelor", "Master", "Professional Degree", "PHD"))
# We group categories 41 and 42 into "Associate"
#table(df$Edu)

### Race ###
#table(df$Race)
df$Race <- ifelse(df$Race >= 6, 6, df$Race)
df$Race <- factor(df$Race, levels = c(1, 2, 3, 4, 5, 6), labels = c("White","Black", "American Indian", "Asian", "Pacific Islander", "Mixed"))
#table(df$Race)

### Disabled ###
df$Pwd <- factor(df$Pwd, levels = c(0, 1, 2), labels = c("NIU", "Yes", "No"))
#table(df$Pwd)

### Health Status ###
df$Health <- factor(df$Health, levels = c(1, 2, 3, 4, 5), labels = c("Excellent", "Very good", "Good", "Fair", "Poor"))
#table(df$Health)

### Insurance Coverage ###
df$Ins <- factor(df$Ins, levels = c(1, 2), labels = c("Yes", "No"))


### State ###
numeric_codes <- c(1, 2, 4, 5, 6, 8, 9, 10, 11, 12, 13, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 44, 45, 46, 47, 48, 49, 50, 51, 53, 54, 55, 56)

state_names <- c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware", "District of Columbia", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virgnia", "Washington", "West Virginia", "Wisconsin", "Wyoming")

df$State <- factor(df$State, levels = numeric_codes, labels = state_names)

# Finally, we update the theme of ggplot to center the titles by default:
theme_update(plot.title = element_text(hjust = 0.5))
```


## A first overview

The overall perception of self state of health of people living in the U.S. based on the Annual Social and Economic Supplement (CPS-ASEC) 2023 is as follows:
```{r}
library("ggplot2")
# Perception of Self State of Health in the United States
ggplot(df, mapping=aes(x=fct_rev(Health), weight = Wt/1e6))+
  geom_bar(fill="#1F78B4", alpha=.6, width=.6) + 
  geom_text(stat = "count", aes(label=sprintf("%.1fM", ..count..)), vjust = -0.2, hjust = -0.1, color = "#004080", size = 3, fontface = "bold" ) +   
geom_text(stat = "count", aes(label = sprintf("(%.1f%%)", ..count.. / sum(..count..) * 100)),
            vjust = 1.3, hjust = -0.1, color = "#1F78B4", size = 3, fontface = "bold") +
  coord_flip() +
  scale_y_continuous(labels = scales::comma, limits = c(0, 120)) +
  labs(x = "Health Status",y = "Weighted Frequency (in millions)", caption = "Source: own elaboration for EDAV Final Project based on CPS-ASEC 2023.") + 
  ggtitle("Perception of Self State of Health in the United States") + 
  theme(panel.background = element_rect(fill = "white"))
```
As it was explained on the Data Section, the CPS-ASEC 2023 contains information on 146,133 individuals, so to build the graph for the entire population we use the variable that weights each observation.

The health status category that is most reported is "Very good" with 33.2% of the population, followed by "Excellent" with a share of 28.9% and "Good" with 27%. Combined together, almost 9 out of 10 people report an at least good health status. 

On the contrary, 9 million people (2.7%) suffer a poor health status and 27 million people (8.2%) a fair one.

In the next sections we will study how the self state of health correlates with other dimensions such as employment, earnings, race, education, among others.

## An expected relationship: Health and Age

In the following graph we display the distribution of age for each state of health: 

```{r}
ggplot(df, aes(Age)) +  
  geom_histogram(aes(y = after_stat(density)), color = "blue", fill = "lightblue", bins = 9, boundary = 0) +  
  facet_grid(~ Health, scales = "free") +  
  labs(caption = "Source: own elaboration for EDAV Final Project based on CPS-ASEC 2023.") + 
  ggtitle("Relationship between Age and Health in the United States") + 
  theme(legend.title = element_text(size = 8)) +
  theme_minimal()
```

As it could be expected, health status worsens with age. For those people who enjoy an excellent health status, the bin with the highest density is the first one, i.e., the bin that includes the youngest people. In addition, the distribution is right-skewed. On the contrary, people who suffer a weak state of health (those with fair or poor health) tend to be older. Indeed, the bins with the highest density are around 70 years old and the distributions are left-skewed. 

We have confirmed something that is quite reasonable: young people are healthier than older people.

In the next sections we are going to study how health relates with a variety of dimensions, such as race, income, employment, among others.

## Association between Health and dimensions Race, Sex and Employment

On this section we study how health relates with Race, Sex and Employement using mosaic plots.

```{r}
library("grid")
library("forcats")

df_1 <- filter(df, (Race == "White" | Race == "Black" | Race == "Asian") & Empt == "Employed")
df_1 <- df_1 %>% mutate(Health = fct_rev(Health))
df_1$Race <- droplevels(df_1$Race)

vcd::mosaic(
  Health ~ Sex + Race, 
  data = df_1, 
  direction = c("v", "v", "h"), 
  weight = df_1$Wt,  
  highlighting_fill = c("red", "orange", "grey90", "lightgreen","darkgreen"),
  labeling = vcd::labeling_border(
    rot_labels = c(0, 0, 45, 0), 
    gp_labels = gpar(fontsize = 7),  
    gp_varnames = gpar(fontsize = 9, fontface = 2), 
    just_labels = c("center", "right"), 
    rot_varnames = c(0, 0, 0, 0)
  ),
  spacing = vcd::spacing_equal(sp = unit(0.3, "lines")),
  main = "Association for Employed People"
)
```


```{r}
df_2 <- filter(df, (Race == "White" | Race == "Black" | Race == "Asian") & Empt == "Unemployed")
df_2 <- df_2 %>% mutate(Health = fct_rev(Health))
df_2$Race <- droplevels(df_2$Race)

vcd::mosaic(
  Health ~ Sex + Race, 
  data = df_2, 
  direction = c("v", "v", "h"), 
  weight = df_1$Wt,  
  highlighting_fill = c("red", "orange", "grey90", "lightgreen","darkgreen"),
  labeling = vcd::labeling_border(
    rot_labels = c(0, 0, 45, 0), 
    gp_labels = gpar(fontsize = 7),  
    gp_varnames = gpar(fontsize = 9, fontface = 2), 
    just_labels = c("center", "right"), 
    rot_varnames = c(0, 0, 0, 0)
  ),
  spacing = vcd::spacing_equal(sp = unit(0.3, "lines")),
  main = "Association for Unemployed People"
)
```

First, it is worth noting that we have drawn different graphs for employed and unemployed people, since they have different sizes.

For both the employed and unemployed people, we see that the proportion of women that suffers a weak state of health is greater than the proportion of men. For employed people, it is specially more notorious for those that has fair health (i.e., the regions shaded in orange), whereas for unemployed people it is clear for both that suffer a fair or poor health (i.e., the regions shaded in orange and red). 

Regarding the race, for both employed and unemployed people, the proportion of black people that suffers a weak state of health is greater than the proportion corresponding to white or asian people. Particularly, Asian population seems to have a quite good state of health (for unemployed ones, even better than the health of white people).

Comparing the two graphs, i.e. the population of employed and unemployed people, we see that for all races and sex, the employed people have a better health than the unemployed people which makes sense. In addition, the group of people that has the highest proportion of people with weak health (i.e., fair or poor health) is the black women without employment showing inequalities on health across U.S. people in terms of sex, race and employment status.

In this section we have learned that there appears to be a difference in health status depending on whether a person is currently employed or unemployed. In the next section we focus on those who are employed and study whether there is a difference on health based on earnings.

## How does earnings associate with health?

In this section we analyze how the distribution of earnings of employed people differs based on the health status. 

```{r}
library(ggridges)
filter(df, Empt == "Employed" & Age>=18) %>%
ggplot(df, mapping=aes(x = Earn, y = fct_rev(Health)), weight = Wt/1e6) +
  geom_density_ridges(fill = "blue",alpha = .5, scale = 0.9) +
  geom_boxplot(pch = 21, width = 0.5, alpha = 0.3) + 
  ggtitle("Distribution of Earnings for Employed People by Health Status") +
  labs(y = "Health", x = "Earnings (in millions)") + 
  scale_x_continuous(labels = scales::comma_format(scale = 1e-6),limits = c(0, 1200000)) 

# The graph would be the same if we order y-axis by the median of earnings y= reorder(Health, Earn, median).
```

The previous graph is quite interesting. First, it shows that the distribution of earnings is right-skewed for every category of health status. In addition, it shows that there seems to be a positive correlation between earnings and health. Indeed, looking at the boxplots, it follows that the both the first quartile (Q1), the median and the third quartile (Q3) of earnings increases with health status. Indeed, for instance the median of people that have an "excellent" -perceived- state of health is higher than the mean of those that have a "very good" state of health, which is in turn higher than the median corresponding to people that have a "good" state of health and so on. In other words, wealthier people enjoys a better state of health than poorer people.

## Are more educated people healthier?

In this section we study whether or not more educated people are healthier. In particular, we analyze the different shares of educational attainment for each level of health status.

In the first part we split the population (aged 18 or more) by citizenship, and in the second part by race. 

### Analysis by Citizenship

The following graph shows the shares of the population with different educational attainment for the each level of health status, by citizenship.
```{r}
df <- df %>%
  mutate(Edu_Grouped = case_when(
    Edu %in% c("College Inc.", "Associate") ~ "College Inc/Assoc.",
    Edu %in% c("Bachelor", "Master", "Professional Degree", "PHD") ~ "Adv. Edu",
    TRUE ~ Edu
  ))

df_3 <- filter(df, Age>=18) %>%
  group_by(Citz, Health, Edu_Grouped) %>%
  dplyr::summarize(Freq = sum(Wt))

df_3 <- df_3 %>%
  group_by(Citz, Health) %>%
  mutate(Total = sum(Freq)) %>%
  ungroup()

ggplot(df_3, aes(x = Edu_Grouped, y = Health)) +
  geom_tile(aes(fill = (Freq/Total)), color = "white") +
  coord_fixed() + 
  scale_fill_gradient2(low = "black", mid = "white",
                        high = "green", midpoint = .2, name = "Share") +
  facet_wrap(~Citz) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0)) +
  labs(x = "Educational Attainment", caption = "Source: own elaboration for EDAV Final Project based on CPS-ASEC 2023.\n Note: 'Adv. Edu' stands for Advanced Education and includes Bachelor degree or higher.") + 
  ggtitle("Share of Educational Attainment by Health Status and Citizenship in the U.S.")
```

Let us analyze first the graph corresponding to Native U.S. citizens. 

We see that for those people who enjoy an excellent or very good state of health, the square corresponding to the 'Adv edu' (which stands for 'Advanced Education', i.e., bachelor degree or higher) is filled in green color. This implies that it is quite probable for a person who enjoys a really good health status to have an advanced education degree. On the contrary, we can see that for these groups of people the squares associated with 'H.S. Inc' (High School Incomplete) are dark grey. This means that it is quite unlikely that the healthiest people have not completed at least high school. 

On the other hand, if we analyze people who suffer a weak state of health (i.e., a poor or fair one) we see that the squares corresponding to advanced education are 'not longer' green, instead they are grey and white. This shows that for people with weak health it is quite unlikely to have an advanced level of education. On the contrary, as it can be seen on the graph, it is quite probable to 'only' have reached a high school degree. 

This positive correlation between health and educational attainment can be also understood in a different way. For the columns corresponding to higher levels of education (i.e., advanced education), the colors range from darker grey colors at the top (poor health) to green colors at the bottom (excellent health), whereas something opposite occurs for more basic levels of education.

Something similar happens to the other citizenship categories such as "Naturalized" or "Not U.S. Citizen". It is particularly interesting to note that for those non U.S. citizens that suffer a weak state of health (i.e., poor or fair) it is quite likely to have a poor level of education since the corresponding squares are green. 

### Analysis by Race

```{r}
df_4 <- filter(df, Age>=18 & (Race == "White" | Race == "Black" | Race == "Asian")) %>%
  group_by(Race, Health, Edu_Grouped) %>%
  dplyr::summarize(Freq = sum(Wt))

df_4 <- df_4 %>%
  group_by(Race, Health) %>%
  mutate(Total = sum(Freq)) %>%
  ungroup()

ggplot(df_4, aes(x = Edu_Grouped, y = Health)) +
  geom_tile(aes(fill = (Freq/Total)), color = "white") +
  coord_fixed() + 
  scale_fill_gradient2(low = "black", mid = "white",
                        high = "green", midpoint = .2, name = "Share") +
  facet_wrap(~Race) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0)) +
  labs(x = "Educational Attainment", caption = "Source: own elaboration for EDAV Final Project based on CPS-ASEC 2023.\n Note: 'Adv. Edu' stands for Advanced Education and includes Bachelor degree or higher.") + 
  ggtitle("Share of Educational Attainment by Health Status and Race in the U.S.")
```

When we split the population by race instead of citizenship, we continue observing an positive correlation between health and education.

For instance, we see that for those people who enjoy a really good state of health, the square corresponding to advanced education is filled in green color (especially for white and asian people). This implies that it is quite probable for a person who enjoys a really good health status to have an advanced education degree. On the contrary, we can see that for these groups of people the squares associated with a low education level are dark grey. This means that it is quite unlikely that the healthiest people have not completed at least high school. 

## Expenditure on Health 

On the following graph we study how the expenditure on health changes across the different health status and over income.

```{r}
set.seed(3267)
samp_idx <- sample(seq_len(nrow(df)), size=400, prob=df$Wt)
new_data <- df[samp_idx, ]

p <- ggplot(new_data, aes(x = Inc, y = MExp, label = State)) +
  geom_point(aes(color = Ins), size = 0.8, alpha = 0.5) +
  scale_color_manual(values = c("Yes" = "green", "No" = "red")) +
  facet_wrap(~Health, scales = "free") + 
  scale_x_continuous(labels = scales::comma_format(scale = 1e-3)) +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3)) +
  labs(x = "Income (in thousands)",y = "Expenditure on health (in thousands)", caption = "Source: own elaboration for EDAV Final Project based on CPS-ASEC 2023.") + 
  labs(color = "Insurance") +
  ggtitle("Expenditure on Health") + 
  theme(legend.title = element_text(size = 8)) +
  theme_minimal() 

p_plotly <- plotly::ggplotly(p)

p_plotly
```
We have chosen a sample of 400 individuals, taking into account the weight of each observation in the entire population. Therefore, as we have explained with the first graph of this section and can be seen in this new graph, the number of people in the sample with very good health is the highest and the number of people with poor health is the smallest. 

The graph shows that there seems to be a positive correlation between the income and the expenditure on health, which makes sense. Indeed, for every state of health, the points appear to increase in both directions, that is, the higher the income, the higher the health expenditure.

Also, we see that people who enjoy a better state of health tend to be richer and also to spend more money on health, which can be explained if they acquire for example better insurance. It can be noted from the graph looking at the scale of both axis for the different facets. For instance, we can see that there are a bunch of people who enjoy an excellent or very good state of health that have an income greater than 100 thousand dollars, however this is something really unlikely for people who suffer a weak state of health. Something similar happens in terms of the expenditure on health.

Regarding the insurance coverage, we see that most people in the U.S. have insurance coverage, regardless of their health status.

## A deeper analysis through an Interactive Parallel Coordinates Plot

Out dataset contains three continuous variables: Inc (income), Earn (earnings) and MExp (medical expenditures). In the following graph we draw an Interactive Parallel Coordinates Plot using these three variables and adding the categorical variables State and Health.

```{r}
df_6 <- df %>% 
  filter(Age >=18) %>%
  select(State,Health,Inc,Earn,MExp,Wt)

df_6$Health <- factor(df_6$Health, levels = c("Excellent", "Very good", "Good", "Fair", "Poor"), labels = c("5-Excellent", "4-Very good", "3-Good", "2-Fair", "1-Poor"))
  
set.seed(3203)
samp_idx <- sample(seq_len(nrow(df_6)), size=150, prob=df_6$Wt)
new_data <- df_6[samp_idx, ]

library(parcoords)
new_data %>% 
  select(-Wt) %>% 
  parcoords(rownames = F,
            brushMode = "1d-multi",
            color = list(colorBy = "Health",
                         colorScale = "scaleOrdinal",
                         colorScheme = "schemeCategory10"),
            alpha = 0.5,
            withD3 = TRUE,
            width = 770,
            height = 600)
```

We have chosen a sample of 150 individuals, taking into account the weight of each observation in the entire population. We subset the sample to include only people aged 18 years or older.

We have added the variable Health so that we can interactively see how the plot changes when we select only a certain health status. Unfortunately, it is not possible for this graph to choose the color assigned to each category.

In the previous graph we can see that if we select the individuals who enjoy an excellent or very good state of health they usually have a higher income, higher earnings and they spend a higher amount on medical expenses (probably because they acquire better insurance coverage) than people who suffer a poor or fair state of health.

We can also see on the graph that the lines crossing from column "Inc" (income) to column "Earn" (Earnings) show that in general higher incomes are associated with higher earnings. Similarly, the lines crossing from earnings to medical expenditure also suggest a positive correlation (i.e., greater the earnings, greater the money spent on health-related issues).

## How Health Status differs across States?

In this section we study how the distribution of health status changes across states. It will allow us to identify those states that face the biggest challenges in terms of health.
```{r}
df_5 <- df %>%
  group_by(State, Health) %>%
  dplyr::summarize(Freq = sum(Wt))

df_5 <- df_5 %>%
  group_by(State) %>%
  mutate(Total = sum(Freq)) %>%
  ungroup()
  
df_5 <- df_5 %>%
  mutate(Health_Share = Freq / Total) %>%
  ungroup()

df_5_sum <- df_5 %>%
  group_by(State) %>%
  summarise(Sum_Poor_Fair = sum(Health_Share[Health %in% c("Poor", "Fair")]))

# Order the levels of "State" based on the sum
df_5$State <- factor(df_5$State, levels =  df_5_sum$State[order(df_5_sum$Sum_Poor_Fair)])

p <- ggplot(df_5, aes(x = State, y = Health_Share, fill = Health)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Excellent" = "darkgreen", "Very good" = "lightgreen","Good" = "grey", "Fair" = "orange", "Poor" = "red")) +
  coord_flip() +
  labs(x = "State",y = "Share by Health Status", caption = "Source: own elaboration for EDAV Final Project based on CPS-ASEC 2023.") + 
  ggtitle("Ranking of States based on Health Status") + 
  theme(axis.text.x = element_text(size = 7), axis.text.y = element_text(size = 7)) +
  theme_minimal()

p_plotly <- plotly::ggplotly(p)

p_plotly
```

On the above interactive plot, we see the heterogeneity of health status across states in the United States. The states are ordered in terms of the proportion of the population that suffer a weak state of health (poor or fair). To build the graph we have taken into account the weight of each observation in the overall population.

From the graph, it follows that Mississippi has the highest proportion of  population suffering a weak health status compared to other states. Indeed, around 18% of their population have a fair (13.7%) or poor (4.4%) state of health. It is followed by Kentucky, West Virginia, Arkansas and New Mexico. In particular, Kentucky is the state with the highest proportion of population that suffer a poor state of health (i.e., the region filled in red is the longest). 

On contrary, there are some states that have a really good performance, for example Utah, South Dakota, Connecticut, Idaho, New Hampshire and Colorado. Utah is the state with the lowest proportion of people suffering a weak (fair or poor) state of health: only 5.7% in total. Particularly, it is explained almost exclusively by the fair state of health since the most extreme category (poor state of health) only accounts for less than 1%. 

The case of New Hampshire is particularly interesting since the share of population that enjoy a really good state of health (very good or excellent) is the highest. Indeed, these two categories account together for more than 75% of the population.

### How do our results change when considering Race?

In this section we deepen on the distribution across states considering also the Race. 

```{r}
df_5 <- df %>%
  filter(Race %in% c("White", "Black", "Asian")) %>%
  group_by(State, Health, Race) %>%
  dplyr::summarize(Freq = sum(Wt))

df_5 <- df_5 %>%
  group_by(State, Race) %>%
  mutate(Total = sum(Freq)) %>%
  ungroup()
  
df_5 <- df_5 %>%
  mutate(Health_Share = Freq / Total) %>%
  ungroup()

df_5_sum_White <- df_5 %>%
  filter(Race == "White") %>%
  group_by(State) %>%
  summarise(Sum_Poor_Fair = sum(Health_Share[Health %in% c("Poor", "Fair")]))

# Order the levels of "State" based on the sum
df_5$State <- factor(df_5$State, levels =  df_5_sum_White$State[order(df_5_sum_White$Sum_Poor_Fair)])

p <- ggplot(df_5, aes(x = State, y = Health_Share, fill = Health)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Excellent" = "darkgreen", "Very good" = "lightgreen","Good" = "grey", "Fair" = "orange", "Poor" = "red")) +
  coord_flip() +
  labs(x = "State",y = "Share by Health Status", caption = "Source: own elaboration for EDAV Final Project based on CPS-ASEC 2023.") + 
  ggtitle("Ranking of States based on Health Status by Race") + 
  theme(axis.text.x = element_text(size = 7), axis.text.y = element_text(size = 7)) +
  facet_wrap(~Race, scales = "free_x") +
  theme_minimal()

p_plotly <- plotly::ggplotly(p)

p_plotly
```

First, on the above interactive graph the states are ordered in terms of the proportion of the population of White people that suffer a weak state of health (poor or fair).

It is worth noting that the distribution of health status across states for White race is similar to the aggregated distribution we discussed in the previous section. It makes sense since the White race is the most frequent in the U.S. population. For instance, four of the 5 states with the lowest ranking are the same for the entire population and for the population of white people: Mississippi, Kentucky, West Virginia and Arkansas.

From the graph we see that, for almost every state, the portion of the population that suffers a weak state of health (i.e., fair or poor) is greater for the black race than for the white race. For instance, for DC less than 4% of white people suffer a weak state of health, but this number rises to around 15% for black people. The case of New Hampshire is weird since more than 50% of the black population suffer a weak state of health. Looking at the raw data it follows that it is due to a small sample size.

Regarding the asian community, they appear to have an overall good state of health, even better than white people in many states. However, its heterogeneity also appears to be bigger (which makes sense due to a small sample size compared to white people).
